---
title: "DATA_605_Assignment_11_DThonn"
file: DThonn_Assign11_DATA_605
name: Daniel Thonn
output:
  html_document: default
  pdf_document: default
course: DATA 605 CUNY MSDA program
---

** DATA_605_Assignment_11_Thonn - Linear Regression-1  **

http://rpubs.com/danthonn/380010

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

``````{r, echo=TRUE}
# install libraries if needed
#install.packages("permutations")
library(permutations)

#install.packages('gtools')
library(gtools)

#install.packages('gvlma')
library(gvlma)

#install.packages('lmtest')
library(lmtest)

```

** Assignment HW 11 **

Using the "cars" dataset in R, build a linear model for stopping distance as a function of speed and replicate the analysis of your textbook chapter 3 (visualization, quality evaluation of the model, and residual analysis.)

``````{r, echo=TRUE}

# load R dataset cars into a dataframe
cars1 <- cars
head(cars1)
#print(cars1)

# examine the dataframe cars1 structure
str(cars1)

```

# create a linear model and Visualize the data

``````{r, echo=TRUE}
# create a linear regression model for cars1
cars1.lm <- lm(cars1$dist ~ cars1$speed)

# plot the cars data
plot(cars1$speed,cars1$dist)
abline(cars1.lm)


# Note: the data looks fairly linear and the model fits reasonably well by appearance

```

Check a summary of the cars1.lm model

``````{r, echo=TRUE}
summary(cars1.lm)

# Findings:
# The pr of the model is less than .05 (.0123) so is significant
# The equation for the line is (y=stop_distance) y = -17.5791 + 3.9324 * x (x=speed)


```

Check the residuals
``````{r, echo=TRUE}

hist(cars1.lm$residuals)
# the histogram of residuals shows a partial shape of normal distribution though skewed to the right
# so not a perfect fit

plot(fitted(cars1.lm),resid(cars1.lm))
# the residuals are randomly distributedon on the residual chart which indicates good linearity

# the sample model fits the theoretical between -2 and +1.5, thought there is divergence after +1.5
# therefore, the model is a partial fit with some divergence
qqnorm(resid(cars1.lm))
qqline(resid(cars1.lm))


```



Check the gvlma function for overall criteria

``````{r, echo=TRUE}
gvlma(cars1.lm)
# Note: there are several issues found in the output of this test.  See below for details.

# Conclusion: 
# There are issues in the gvlma check as shown below with prob < .05 for (3) criteria.
# Heteroscedasticity, Skewness, and Global Stat fail
# Conclusion: the cars1.lm model is less than an ideal model for the cars data.

# Global Stat        15.801 0.003298 Assumptions NOT satisfied!
# Skewness            6.528 0.010621 Assumptions NOT satisfied!
# Kurtosis            1.661 0.197449    Assumptions acceptable.
# Link Function       2.329 0.126998    Assumptions acceptable.
# Heteroscedasticity  5.283 0.021530 Assumptions NOT satisfied!

```




** END **

``````{r, echo=TRUE}

```




